### 2023-04-26 chapter0/chapter0.tex
Tweaked sentence about broader impacts of study

These analyses were conducted using IDRs identified in various species of yeast, which is a widely used model organism in molecular biology research. However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would indicate the existence of a taxonomy of IDRs which is conserved across the tree of life.

>>>>

These analyses were conducted using IDRs identified in various species of yeast, which is a widely used model organism in molecular biology research. However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in gene regulation as well as human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would indicate the existence of a taxonomy of IDRs which is conserved across the tree of life.

### 2023-04-26 chapter2/chapter2.tex
Tweaked sentence about broader impacts of study

However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would indicate the existence of a taxonomy of IDRs which is conserved across the tree of life.

% III. Many of these analyses were limited to ... However, a recent study first to demonstrate patterns of selection in IDRs across an entire proteome using a set of 82 ``molecular features''
%    D. Furthermore, by clustering IDRs on by these ``evolutionary signatures,'' they identified groups of IDRs evolving under similar sets of constraints. Furthermore they identified functional associations in these clusters which can be used to generate hypotheses of function for previously uncharacterized proteins
%    E. Though subsequent studies have expanded on these findings by using techniques from machine learning and statistics the information from evolutionary signatures to predict the function of IDRs. and learn disorder-associated features in a self-supervised manner.
% III. However, no subsequent studies have determined if similar patterns of conservation are found in other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}.
%    A. Expand analyses to another model organism
%        1. Drosophila is key model for genetic and development research with abundant genomic resources across many lineages in the radiation
%        2. Furthermore as multicellular organism with complex developmental pathways, the results would be particularly relevant for understanding the role of IDRs in complex signaling pathways
%        3. Also relevant for human health!
%    B. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. Depending on the amount of concordance, several conclusions are possible.
%        1. If no patterns are observed, this suggests that conservation of distributed is a not widely conserved mechanism of IDR evolution.
%        2. Patterns of conservation are yeast phenomena
%        3. Finally, the identification of similar clusters with conserved features and properties would suggest the existence of ``taxonomy'' of IDRs which is common across many branches of the tree of life. This would allow the creation of a classification system such as Pfam
%
% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework through the application of the Brownian motion and Ornstein-Uhlenbeck models of trait evolution.
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
%
% V. Therefore, in this work we apply evolutionary models of sequence and trait evolution to the sequences of IDRs and their of associated molecular feature to dissect the forces driving their evolution. We find IDRs exhibit unique patterns of evolution at the level of sequences. We find intrinsic disorder is a dynamically evolving property in some regions, and changes in disorder preference are correlated with changes in the complexity. We find that IDRs are unconstrained along several axis of trait evolution, but they also display

>>>>

However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in gene regulation as well as human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would indicate the existence of a taxonomy of IDRs which is conserved across the tree of life.
% The latter result would represent a significant step towards the creation of resources for the classification of IDRs analogous to those for folded domains such as Pfam, CATH or SCOP.

% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework through the application of the Brownian motion and Ornstein-Uhlenbeck models of trait evolution.
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
%
% V. Therefore, in this work we apply evolutionary models of sequence and trait evolution to the sequences of IDRs and their of associated molecular feature to dissect the forces driving their evolution. We find IDRs exhibit unique patterns of evolution at the level of sequences. We find intrinsic disorder is a dynamically evolving property in some regions, and changes in disorder preference are correlated with changes in the complexity. We find that IDRs are unconstrained along several axis of trait evolution, but they also display
